Synthesis of Linear Functional Programs

### 2021-05-09 article/article.tex

Even with everything mentioned so far, non-determinism is still
present proof-search in logic, e,g., at any given point, many proof
rules are applicable in general. The technique of focusing~\cite{10.1093/logcom/2.3.297}
has been previously studied as a way to discipline proof search in
linear logic-- a method created to trim down the search space of
valid proofs in linear logic, by eagerly applying invertible rules
(i.e.~rules whose conclusion implies the premises), and then by
``focusing'' on a single connective when no more direct (invertible)
rules can be applied, that is, only applying rules that breakdown
the connective under focus or its subformulas. If the search is not
successful, the procedure backtracks and another connective is
chosen as the focus.

@@

\mypara{Step by step} Our work so far has been constantly accompanied
by somewhat dense theoretical material, state of the art,
which we must understand to overcome challenges in these fields,
and build from the start a more scalable synthetizer -- we want
to display concrete results to prove the feasability of synthetising
linear types, and thus possibility for future application to related fields,
such as session types for message passing concurrency.

Apart from the theory, when we started studying type systems, we created
a simple type-checker for the simply-typed $\lambda$-calculus.
While learning about linear type systems, and the simply-typed linear
$\lambda$-calculus, we developed a parser, a desugar module, and an
evaluation module to write simple programs in the language;
culminating in writing the type-checker for linear types. At this point,
where we stand as of now, we can parse, type-check, and evaluate this
small linear functional language.
The next step is to start developing the actual synthetizer, embedding it
with all the techniques we've seen. Following work will be done to
augment usability, expand the language domain, and add features such as the
ones mentioned above.

>>>>

Even with everything mentioned so far, non-determinism is still
present proof-search in logic, e,g., at any given point, many proof
rules are applicable in general. The technique of focusing~\cite{10.1093/logcom/2.3.297,DBLP:conf/cade/ChaudhuriP05}
has been previously studied as a way to discipline proof search in
linear logic -- a method created to trim down the search space of
valid proofs in linear logic, by eagerly applying invertible rules
(i.e.~rules whose conclusion implies the premises), and then by
``focusing'' on a single connective when no more direct (invertible)
rules can be applied, that is, only applying rules that breakdown
the connective under focus or its subformulas. If the search is not
successful, the procedure backtracks and another connective is
chosen as the focus.

@@

\mypara{Step by step} Our work so far has been constantly accompanied
by the study of somewhat dense theoretical material, state of the art,
which we must understand to overcome challenges in these fields, in
order to develop a more scalable synthetizer -- we want to display
concrete results to prove the feasability of program synthesis from
linear types, and thus open the possibility for future application to
related fields, such as program synthesis based on session types for
message-passing concurrency.

Apart from the theory, when we started studying type systems, we
created a simple type-checker for the simply-typed $\lambda$-calculus.
While learning about linear type systems, and the simply-typed linear
$\lambda$-calculus, we developed a parser, a desugaring module, and an
evaluation module to write, analyze and execute simple programs in the
language; culminating in writing a type-checker for the linear
$\lambda$-calculus. At this point, we can parse, typecheck, and
execute this small linear functional language.  The next step is to
start developing the actual synthetizer, embedding it with all the
techniques we've seen. Following work will be done to augment
usability, expand the language domain, and add features such as those
mentioned above.
